---
title: ""
output:
  html_document:
    theme: null
    self_contained: false
    mathjax: null
    highlight: null
    toc: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
esc <- function(x) as.character(htmltools::htmlEscape(as.character(x)))
num <- function(x) ifelse(is.na(x), "Not available", format(x, trim = TRUE, scientific = FALSE))
numeric_span <- function(x) {
  value <- esc(num(x))
  paste0('<span data-expected="', value, '">', value, '</span>')
}
table_number <- 0L
figure_number <- 0L
report_table <- function(data, caption, numeric_columns = character()) {
  table_number <<- table_number + 1L
  headers <- paste0("<th scope='col'>", esc(names(data)), "</th>", collapse = "")
  rows <- vapply(seq_len(nrow(data)), function(i) {
    cells <- vapply(names(data), function(k) {
      value <- esc(data[[k]][i])
      if (k %in% numeric_columns) paste0('<td class="numeric" data-expected="', value, '">', value, '</td>')
      else paste0("<td>", value, "</td>")
    }, "")
    paste0("<tr>", paste(cells, collapse = ""), "</tr>")
  }, "")
  cat(paste0("<table><caption>Table ", table_number, ". ", esc(caption),
    "</caption><thead><tr>", headers, "</tr></thead><tbody>", paste(rows, collapse = ""), "</tbody></table>\n"))
}
mode_sentence <- function(v) {
  a <- report$items[[v]]
  if (a$summary$Percent_denominator == 0L) return("No mapped valid responses were available.")
  t <- a$table
  top <- which(t$Count == max(t$Count))
  labels <- paste0("‘", esc(t$Response_option[top]), "’", collapse = " and ")
  tied <- length(top) > 1L
  paste0(if (tied) "The joint most frequent categories were " else "The most frequent category was ",
    labels, " (", numeric_span(t$Count[top[1]]),
    if (tied) " responses per category; " else " responses; ",
    numeric_span(sprintf("%.1f", t$Percent_of_valid_responses[top[1]])),
    if (tied) "% each, among " else "% of ",
    numeric_span(a$summary$Percent_denominator), " valid responses).")
}
item_results <- function(v) {
  a <- report$items[[v]]; s <- a$summary
  cat('<div class="item"><h3>', esc(a$title), '</h3>\n', sep = "")
  cat('<p class="denominator">Percentages: valid mapped n = ', numeric_span(s$Percent_denominator),
    '; total = ', numeric_span(s$Total_responses), '; missing = ', numeric_span(s$Missing_responses),
    '; unmapped = ', numeric_span(s$Unmapped_nonmissing_responses), '.</p>\n', sep = "")
  t <- data.frame(Order = a$table$Questionnaire_order, `Response option` = a$table$Response_option,
    Count = a$table$Count, `Valid responses (%)` = ifelse(is.na(a$table$Percent_of_valid_responses),
      "Not available", sprintf("%.1f", a$table$Percent_of_valid_responses)), check.names = FALSE)
  report_table(t, paste0(v, ": response distribution."),
    c("Order", "Count", "Valid responses (%)"))
  # A modal-category sentence adds no information beyond the table and chart.
  # Keep these concise sentences in the overview rather than repeating each item.
  unknown <- sum(is.na(a$definitions$Confirmed_exported_code))
  if (unknown > 0L) cat('<p class="footnote">Mapping review: ', numeric_span(unknown),
    ' option(s) have unconfirmed export codes.</p>\n', sep = "")
  if (nrow(a$unmapped)) report_table(a$unmapped, paste("Unmapped values requiring review:", v), "Count")
  if (is.na(a$asset)) {
    cat('</div>\n')
    return(invisible(NULL))
  }
  figure_number <<- figure_number + 1L
  cat('<figure><img src="', esc(a$asset), '" alt="', esc(paste(a$title, "bar chart; all values are in the preceding table.")),
    '"><figcaption>Figure ', figure_number, '. ', esc(v),
    ': counts and valid-response percentages.</figcaption></figure></div>\n', sep = "")
}
age <- report$age
```

```{r title, results='asis'}
cat('<div class="title-card"><div class="eyebrow">EUAqua · Work Package 4</div>',
  '<h1>Italian pilot<br>Preliminary questionnaire report</h1>',
  '<p>Descriptive evidence from the pilot sample</p><div class="metadata">',
  '<div><strong>Pilot date</strong>', esc(format(report$pilot_date, "%d %B %Y")), '</div>',
  '<div><strong>Report date</strong>', esc(format(report$report_date, "%d %B %Y")), '</div>',
  '<div><strong>Sample size</strong>', numeric_span(report$sample_size), ' questionnaire responses</div>',
  '</div></div>', sep = "")
```

## Overview

This report describes the preliminary questionnaire responses in the Italian EUAqua pilot. Distributions refer only to this pilot sample.

```{r overview, results='asis'}
for (v in c("D_fish_freq", "D_diet_type", "E_label_trust", "E_label_format", "E_oa_awareness")) {
  cat('<p><strong>', esc(report$items[[v]]$title), '.</strong> ', mode_sentence(v), '</p>\n', sep = "")
}
cat('<p><strong>Age.</strong> Median ', numeric_span(age$Median_years), ' years; first and third quartiles ',
  numeric_span(age$First_quartile_years), ' and ', numeric_span(age$Third_quartile_years),
  ' years; range ', numeric_span(age$Minimum_years), '–', numeric_span(age$Maximum_years),
  ' years, among ', numeric_span(age$Numerical_summary_denominator), ' finite numerical responses.</p>', sep = "")
```

## Methods and completeness

The report reuses verified aggregates and charts from `scripts/03_describe_qualtrics.R`. Preparation matches Values and Labels exports by ResponseId, checks PID consistency and retains QSF definitions and response order. No identifiers enter the report, no participants are excluded and originals remain unchanged.

The saved **EUAQUA WP4 executive Gantt** states “Pilot scheduled for 30 September 2026”, consistent with the archive designation. This is the scheduled session date, distinct from questionnaire completion dates.

For categorical items, **valid** means a nonmissing value with an unambiguous mapping confirmed against the matched exports. **Missing** means NA or an empty string; other categorical values are not silently trimmed or recoded. Unmapped nonmissing values are counted separately. Each percentage is the category count divided by that item's valid mapped responses, multiplied by 100. Values are displayed to one decimal place; rounding may affect displayed sums.

Tables retain all options in questionnaire order, including zero counts. Unobserved codes remain unassigned unless confirmed by the exports; QSF choice IDs are not assumed to be export values. Zero counts reflect matched exports, not assumed codes. “Never”, “Prefer not to answer”, and “I do not read labels” are valid categories, with non-readers included in the denominator. Ordinal items remain separate, without means, scores, combined scales or category collapsing.

Age summaries retain all finite values from a derived numerical copy. Whitespace-only entries are missing; nonblank nonnumeric/nonfinite entries, values outside the QSF range and noninteger ages are flagged without automatic removal. Quartiles use R `quantile(type = 7)`, with linear interpolation at **h = 1 + (n − 1) × p**. The denominator is the number of finite converted values. The user-approved comparison bands are **18–30, 31–55 and 56+**, in completed years. Noninteger or under-18 values are unassigned in the band table without exclusion from numerical summaries. No demographic cross-tabulations are used.

```{r completeness, results='asis'}
o <- report$completeness
cat('<p>Coverage: ', numeric_span(nrow(o)), ' categorical research items; ',
  numeric_span(sum(!vapply(report$items, function(a) is.na(a$asset), logical(1)))),
  ' charts. Gender, education and occupation are reported as separate marginal tables.</p>', sep = "")
t <- data.frame(Item = o$Variable, Total = o$Total_responses, `Valid mapped` = o$Valid_mapped_responses,
  Missing = o$Missing_responses, Unmapped = o$Unmapped_nonmissing_responses,
  `Percentage denominator` = o$Percent_denominator, check.names = FALSE)
report_table(t, "Categorical completeness and item-specific percentage denominators.",
  names(t)[-1])
```

Free-text checks account for whitespace, Unicode spaces and invisible zero-width/BOM characters; no quotations or thematic coding are included. Consent is audited separately. Permissions for subsequent transcript or other analyses require review; this report does not establish blanket authorization.

## Sample characteristics

```{r comparison, results='asis'}
context <- report$context
observed <- function(n, pct) ifelse(is.na(pct), paste0(n, " (percentage unavailable)"),
  sprintf("%d (%.1f%%)", n, pct))
row <- function(category, benchmark, target, n, pct, note) data.frame(
  `Variable/category` = category, `ISTAT benchmark` = benchmark,
  `Recruitment target` = target, `Observed pilot n (%)` = observed(n, pct),
  `Comparability notes` = note, check.names = FALSE)
gender <- report$items$C_gender$table
g <- match(names(context$sex_benchmark), gender$Response_option)
stopifnot(!anyNA(g))
comparison <- row(paste("Sex/gender:", gender$Response_option[g]),
  sprintf("%.2f%% *", context$sex_benchmark), as.character(context$sex_target),
  gender$Count[g], gender$Percent_of_valid_responses[g], "S1")
bands <- report$age_bands
comparison <- rbind(comparison, row(paste("Age:", bands$Category),
  sprintf("%.2f%% *", context$age_benchmark[bands$Category]),
  as.character(context$age_target[bands$Category]), bands$Count, bands$Percent, "A1"))
edu <- report$education_comparison
comparison <- rbind(comparison, row(paste("Education:", edu$Category),
  sprintf("%.1f%%", context$education_benchmark), as.character(context$education_target),
  edu$Count, edu$Percent, "E1"))
fish <- report$items$D_fish_freq$table
comparison <- rbind(comparison, row(paste("Seafood:", fish$Response_option),
  "Not supplied", "Mix; no numerical quota", fish$Count, fish$Percent_of_valid_responses, "F1"))
report_table(comparison, "Contextual benchmarks, planned recruitment targets and observed questionnaire results.",
  c("ISTAT benchmark", "Recruitment target", "Observed pilot n (%)"))
cat('<p class="footnote">* ', esc(context$demographic_status), '. ',
  esc(context$demographic_population), '. The official page supports the date and estimated status; ',
  'the supplied adult percentages and single-year-age aggregation have not been independently reproduced.</p>', sep = "")
cat('<p>', esc(context$statement), '</p>', sep = "")
varying <- sum(vapply(c("C_gender", "C_education", "C_occupation"),
  function(v) sum(report$items[[v]]$table$Count > 0) > 1, logical(1)))
if (varying >= 2L && sum(fish$Count > 0) > 1L)
  cat('<p>The focus group includes variation in demographic characteristics and seafood consumption.</p>')
if (any(bands$Count == 0L)) cat('<p class="footnote">Age coverage review: no observed responses in ',
  esc(paste(bands$Category[bands$Count == 0L], collapse = ", ")),
  '. Variation does not mean that every planned age band is represented.</p>', sep = "")
cat('<p class="denominator">Observed denominators: gender n = ',
  numeric_span(report$items$C_gender$summary$Percent_denominator), '; age bands n = ',
  numeric_span(report$age_band_summary$Valid_banded), '; education n = ',
  numeric_span(report$items$C_education$summary$Percent_denominator), '; seafood frequency n = ',
  numeric_span(report$items$D_fish_freq$summary$Percent_denominator),
  '. Age-band missing = ', numeric_span(report$age_band_summary$Missing),
  '; unassigned nonmissing = ', numeric_span(report$age_band_summary$Unassigned_nonmissing), '.</p>', sep = "")
```

**S1 — Sex/gender.** ISTAT measures registered sex; the questionnaire asks self-reported gender. The female/male recruitment targets are compared descriptively with Woman/Man selections, without treating the measures as equivalent. All questionnaire gender categories remain in the separate distribution and its denominator.

**A1 — Age.** ISTAT context covers adults aged 18+ at 1 January 2026; the observed bands use the questionnaire ages. Targets are a recruitment plan, not observed counts.

**E1 — Education.** ISTAT covers ages 25–64 in 2024, whereas the pilot table covers all valid responses. Its upper-secondary category includes post-secondary non-tertiary qualifications, and tertiary education includes qualifications beyond university degrees. For comparison only, the questionnaire's bachelor's and master's/higher options are combined; all four original options remain below. The recruitment targets concern no qualification below upper secondary, high-school diploma and university degree, respectively. This is an approximate alignment, not an ISCED equivalence or an age-restricted comparison.

**F1 — Seafood.** The recruitment plan sought a mix of frequent and occasional consumers without a numerical quota. No national benchmark was supplied. Observed rows retain actual questionnaire categories; no frequent/occasional classification is constructed.

```{r benchmark_sources, results='asis'}
cat('<p class="footnote">Sources: <a href="', esc(context$demographic_url), '">ISTAT population by sex and age</a>; ',
  '<a href="', esc(context$education_url), '">ISTAT education, 2024</a>; ',
  '<a href="', esc(context$education_pdf), '">full education report</a>. ',
  esc(context$education_status), '</p>', sep = "")
```

```{r age, results='asis'}
fields <- c("Total_responses", "Valid_numeric_responses", "Missing_responses", "Invalid_nonmissing_responses",
  "Median_years", "First_quartile_years", "Third_quartile_years", "Minimum_years", "Maximum_years",
  "Outside_QSF_range", "Noninteger_values", "Numerical_summary_denominator")
labels <- c("Total responses", "Valid finite numerical responses", "Missing responses", "Invalid nonmissing responses",
  "Median (years)", "First quartile (years)", "Third quartile (years)", "Minimum (years)", "Maximum (years)",
  "Outside QSF validation range", "Noninteger values", "Numerical summary denominator")
report_table(data.frame(Measure = labels, Value = vapply(fields, function(k) num(age[[k]]), "")),
  "Age summary and conversion checks.", "Value")
```

```{r demographics, results='asis'}
for (v in c("C_gender", "C_education", "C_occupation")) item_results(v)
```

These are separate questionnaire distributions, including zero-count options, with no demographic cross-tabulations. No occupation benchmark or recruitment quota was supplied. Small marginal groups warrant disclosure review before wider circulation.

## Purchasing and dietary habits

```{r purchasing, results='asis'}
vars <- report$completeness$Variable[report$completeness$Section == "purchasing_and_dietary_habits"]
for (v in vars) item_results(v)
```

## Label and sustainability perceptions

```{r perceptions, results='asis'}
vars <- report$completeness$Variable[report$completeness$Section == "label_and_sustainability_perceptions"]
for (v in vars) item_results(v)
```

## Limitations and unresolved issues

This is a small pilot sample. Its descriptive distributions cannot establish population prevalence, causal relationships or associations between variables. No demographic cross-tabulations, free-text quotations, combined scales or participant exclusions are included. Age extrema are aggregate summaries approved for this step; they should still be reviewed before wider dissemination of a small-sample report.

The education display labels below are **provisional descriptive translations**. They do not assert international qualification equivalence. In particular, the final Italian option combines *laurea magistrale o superiore* with the parenthetical *Master, PhD*; the display label must be reviewed if used beyond this report. Original Italian definitions remain intact in the codebook.

```{r translations, results='asis'}
d <- report$education
report_table(data.frame(`Italian questionnaire option` = d$Original_option,
  `Provisional English display label` = d$English_display_label, check.names = FALSE),
  "Education definitions for translation review; no participant responses or subgroup counts.")
cat('<div class="review"><strong>Typography.</strong> Preferred font: ', esc(report$font$Preferred_font),
  '; installed: ', esc(report$font$Preferred_font_available), '. First fallback: ', esc(report$font$First_fallback),
  '; installed: ', esc(report$font$First_fallback_available), '. The chart renderer resolved regular text to ',
  esc(report$font$Resolved_regular_family), ' and bold text to ', esc(report$font$Resolved_bold_family),
  '. The same regular family is requested for this HTML report. Browser rendering should be rechecked on another computer.</div>', sep = "")
```

The blue, green and light-blue palette comes from the shared EUAqua theme and represents provisional visual approximations, not verified official brand colour codes. No appropriate official logo asset was found in the analysis workspace; no replacement logo has been created and no fonts have been downloaded.

Unconfirmed codes require survey/export evidence before future use. Education translations, exact demographic benchmark calculations and wider disclosure of small marginal groups require review. Correct the pilot-date metadata if the actual session differed from the schedule.

## Consent and permissions

C1–C3 are required in the verified questionnaire participation flow. C4–C5 are optional, activity-specific permissions. Below are English display translations of the verified questions; the exact Italian wording remains in the codebook and consent-definition output.

```{r permissions, results='asis'}
c <- report$consent
defs <- report$consent_definitions
stopifnot(identical(c$Consent_variable, defs$Consent_variable),
  all(c$Granted + c$Declined + c$Missing + c$Unmapped == c$Total_responses))
t <- data.frame(Item = c$Consent_variable,
  Role = ifelse(c$Consent_variable %in% paste0("C", 1:3), "Required in QSF flow", "Optional"),
  `Total / denominator` = c$Total_responses, Agreement = c$Granted, Refusal = c$Declined,
  Missing = c$Missing, Unmapped = c$Unmapped, check.names = FALSE)
report_table(t, "Aggregate consent selections. All four counts use the item's total response denominator.", names(t)[3:7])
for (i in seq_len(nrow(defs))) {
  cat('<p><strong>', esc(defs$Consent_variable[i]), ' — ', esc(defs$Permission[i]),
    '.</strong> ', esc(defs$English_question[i]), '</p>', sep = "")
}
flagged <- c$Consent_variable[c$Review_required]
if (length(flagged)) {
  cat('<div class="review">Permission review flags: ', esc(paste(flagged, collapse = ", ")),
    '. Review the item-specific refusal, missing and unmapped counts before relevant analyses.</div>', sep = "")
} else {
  cat('<p>No refusal, missing or unmapped selections were detected in these exported consent items.</p>')
}
```

C1 concerns participation; C2 concerns personal-data processing for this research; C3 concerns recording and research use of image and voice. C4 specifically concerns AI facial-expression analysis under sections A/A1 of the privacy notice. C5 concerns retention and further use for future research under section B. Optional refusal is a permitted choice, not a participation failure, and does not grant that activity's permission.

Agreement and refusal counts require export-confirmed mappings. Unrecognized nonmissing selections remain unmapped and require review rather than being assigned to either decision.

Before transcript, recording or facial-expression analysis, review the permissions relevant to that activity alongside the referenced privacy notice, approved protocol and applicable requirements. Agreement to one activity does not authorize another, and questionnaire selections alone do not establish every ethical, legal or operational requirement. No individual consent records are displayed and no participants are automatically excluded.

## Reproducibility

In RStudio, run `source("reports/render_pilot_report.R")` from the project root. It reruns the existing analysis, verifies tables against CSV outputs and renders numerical findings directly from aggregates.

Keep the HTML and its asset folders together under ignored `outputs/`. Numerical checks and R environment details are recorded in `verification.txt` and `session_info.txt`. Review browser print preview before printing or saving a PDF.
