---
title: ""
output:
  html_document:
    theme: null
    self_contained: false
    mathjax: null
    highlight: null
    toc: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
esc <- function(x) as.character(htmltools::htmlEscape(as.character(x)))
num <- function(x) ifelse(is.na(x), "Not available", format(x, trim = TRUE, scientific = FALSE))
numeric_span <- function(x) {
  value <- esc(num(x))
  paste0('<span data-expected="', value, '">', value, '</span>')
}
table_number <- 0L
figure_number <- 0L
report_table <- function(data, caption, numeric_columns = character()) {
  table_number <<- table_number + 1L
  headers <- paste0("<th scope='col'>", esc(names(data)), "</th>", collapse = "")
  rows <- vapply(seq_len(nrow(data)), function(i) {
    cells <- vapply(names(data), function(k) {
      value <- esc(data[[k]][i])
      if (k %in% numeric_columns) paste0('<td class="numeric" data-expected="', value, '">', value, '</td>')
      else paste0("<td>", value, "</td>")
    }, "")
    paste0("<tr>", paste(cells, collapse = ""), "</tr>")
  }, "")
  cat(paste0("<table><caption>Table ", table_number, ". ", esc(caption),
    "</caption><thead><tr>", headers, "</tr></thead><tbody>", paste(rows, collapse = ""), "</tbody></table>\n"))
}
mode_sentence <- function(v) {
  a <- report$items[[v]]
  if (a$summary$Percent_denominator == 0L) return("No mapped valid responses were available.")
  t <- a$table
  top <- which(t$Count == max(t$Count))
  labels <- paste0("‘", esc(t$Response_option[top]), "’", collapse = " and ")
  tied <- length(top) > 1L
  paste0(if (tied) "The joint most frequent categories were " else "The most frequent category was ",
    labels, " (", numeric_span(t$Count[top[1]]),
    if (tied) " responses per category; " else " responses; ",
    numeric_span(sprintf("%.1f", t$Percent_of_valid_responses[top[1]])),
    if (tied) "% each, among " else "% of ",
    numeric_span(a$summary$Percent_denominator), " valid responses).")
}
item_results <- function(v) {
  a <- report$items[[v]]; s <- a$summary
  cat('<div class="item"><h3>', esc(a$title), '</h3>\n', sep = "")
  cat('<p class="denominator">Percentages: valid mapped n = ', numeric_span(s$Percent_denominator),
    '; total = ', numeric_span(s$Total_responses), '; missing = ', numeric_span(s$Missing_responses),
    '; unmapped = ', numeric_span(s$Unmapped_nonmissing_responses), '.</p>\n', sep = "")
  t <- data.frame(Order = a$table$Questionnaire_order, `Response option` = a$table$Response_option,
    Count = a$table$Count, `Valid responses (%)` = ifelse(is.na(a$table$Percent_of_valid_responses),
      "Not available", sprintf("%.1f", a$table$Percent_of_valid_responses)), check.names = FALSE)
  report_table(t, paste0(v, ": response distribution."),
    c("Order", "Count", "Valid responses (%)"))
  # A modal-category sentence adds no information beyond the table and chart.
  # Keep these concise sentences in the overview rather than repeating each item.
  unknown <- sum(is.na(a$definitions$Confirmed_exported_code))
  if (unknown > 0L) cat('<p class="footnote">Mapping review: ', numeric_span(unknown),
    ' option(s) have unconfirmed export codes.</p>\n', sep = "")
  if (nrow(a$unmapped)) report_table(a$unmapped, paste("Unmapped values requiring review:", v), "Count")
  figure_number <<- figure_number + 1L
  cat('<figure><img src="', esc(a$asset), '" alt="', esc(paste(a$title, "bar chart; all values are in the preceding table.")),
    '"><figcaption>Figure ', figure_number, '. ', esc(v),
    ': counts and valid-response percentages.</figcaption></figure></div>\n', sep = "")
}
age <- report$age
```

```{r title, results='asis'}
cat('<div class="title-card"><div class="eyebrow">EUAqua · Work Package 4</div>',
  '<h1>Italian pilot<br>Preliminary questionnaire report</h1>',
  '<p>Descriptive evidence from the pilot sample</p><div class="metadata">',
  '<div><strong>Pilot date</strong>', esc(format(report$pilot_date, "%d %B %Y")), '</div>',
  '<div><strong>Report date</strong>', esc(format(report$report_date, "%d %B %Y")), '</div>',
  '<div><strong>Sample size</strong>', numeric_span(report$sample_size), ' questionnaire responses</div>',
  '</div></div>', sep = "")
```

## Overview

This report describes the preliminary questionnaire responses in the Italian EUAqua pilot. Distributions refer only to this pilot sample.

```{r overview, results='asis'}
for (v in c("D_fish_freq", "D_diet_type", "E_label_trust", "E_label_format", "E_oa_awareness")) {
  cat('<p><strong>', esc(report$items[[v]]$title), '.</strong> ', mode_sentence(v), '</p>\n', sep = "")
}
cat('<p><strong>Age.</strong> Median ', numeric_span(age$Median_years), ' years; first and third quartiles ',
  numeric_span(age$First_quartile_years), ' and ', numeric_span(age$Third_quartile_years),
  ' years; range ', numeric_span(age$Minimum_years), '–', numeric_span(age$Maximum_years),
  ' years, among ', numeric_span(age$Numerical_summary_denominator), ' finite numerical responses.</p>', sep = "")
```

## Methods and completeness

The report reuses verified aggregates and charts from `scripts/03_describe_qualtrics.R`. Preparation matches Values and Labels exports by ResponseId, checks PID consistency and retains QSF definitions and response order. No identifiers enter the report, no participants are excluded and originals remain unchanged.

The saved **EUAQUA WP4 executive Gantt** states “Pilot scheduled for 30 September 2026”, consistent with the archive designation. This is the scheduled session date, distinct from questionnaire completion dates.

For categorical items, **valid** means a nonmissing value with an unambiguous mapping confirmed against the matched exports. **Missing** means NA or an empty string; other categorical values are not silently trimmed or recoded. Unmapped nonmissing values are counted separately. Each percentage is the category count divided by that item's valid mapped responses, multiplied by 100. Values are displayed to one decimal place; rounding may affect displayed sums.

Tables retain all options in questionnaire order, including zero counts. Unobserved codes remain unassigned unless confirmed by the exports; QSF choice IDs are not assumed to be export values. Zero counts reflect matched exports, not assumed codes. “Never”, “Prefer not to answer”, and “I do not read labels” are valid categories, with non-readers included in the denominator. Ordinal items remain separate, without means, scores, combined scales or category collapsing.

Age summaries retain all finite values from a derived numerical copy. Whitespace-only entries are missing; nonblank nonnumeric/nonfinite entries, values outside the QSF range and noninteger ages are flagged without automatic removal. Quartiles use R `quantile(type = 7)`, with linear interpolation at **h = 1 + (n − 1) × p**. The denominator is the number of finite converted values. No age bands or demographic combinations are used.

```{r completeness, results='asis'}
o <- report$completeness
cat('<p>Coverage: ', numeric_span(nrow(o)), ' categorical research items; ', numeric_span(length(report$items)),
  ' charts. Gender, education and occupation distributions are withheld; their completeness is retained.</p>', sep = "")
t <- data.frame(Item = o$Variable, Total = o$Total_responses, `Valid mapped` = o$Valid_mapped_responses,
  Missing = o$Missing_responses, Unmapped = o$Unmapped_nonmissing_responses,
  `Percentage denominator` = o$Percent_denominator, check.names = FALSE)
report_table(t, "Categorical completeness. Denominators are item-specific; demographic subgroup counts are not shown.",
  names(t)[-1])
```

Free-text checks account for whitespace, Unicode spaces and invisible zero-width/BOM characters; no quotations or thematic coding are included. Consent is audited separately. Permissions for subsequent transcript or other analyses require review; this report does not establish blanket authorization.

## Sample characteristics

```{r age, results='asis'}
fields <- c("Total_responses", "Valid_numeric_responses", "Missing_responses", "Invalid_nonmissing_responses",
  "Median_years", "First_quartile_years", "Third_quartile_years", "Minimum_years", "Maximum_years",
  "Outside_QSF_range", "Noninteger_values", "Numerical_summary_denominator")
labels <- c("Total responses", "Valid finite numerical responses", "Missing responses", "Invalid nonmissing responses",
  "Median (years)", "First quartile (years)", "Third quartile (years)", "Minimum (years)", "Maximum (years)",
  "Outside QSF validation range", "Noninteger values", "Numerical summary denominator")
report_table(data.frame(Measure = labels, Value = vapply(fields, function(k) num(age[[k]]), "")),
  "Age summary and conversion checks.", "Value")
```

Gender, education and occupation subgroup counts and charts are withheld pending an agreed disclosure rule for this small pilot. This does not collapse categories or exclude participants.

## Purchasing and dietary habits

```{r purchasing, results='asis'}
vars <- report$completeness$Variable[report$completeness$Section == "purchasing_and_dietary_habits"]
for (v in vars) item_results(v)
```

## Label and sustainability perceptions

```{r perceptions, results='asis'}
vars <- report$completeness$Variable[report$completeness$Section == "label_and_sustainability_perceptions"]
for (v in vars) item_results(v)
```

## Limitations and unresolved issues

This is a small pilot sample. Its descriptive distributions cannot establish population prevalence, causal relationships or associations between variables. No demographic cross-tabulations, free-text quotations, combined scales or participant exclusions are included. Age extrema are aggregate summaries approved for this step; they should still be reviewed before wider dissemination of a small-sample report.

The education display labels below are **provisional descriptive translations**. They do not assert international qualification equivalence. In particular, the final Italian option combines *laurea magistrale o superiore* with the parenthetical *Master, PhD*; the display label must be reviewed if used beyond this report. Original Italian definitions remain intact in the codebook.

```{r translations, results='asis'}
d <- report$education
report_table(data.frame(`Italian questionnaire option` = d$Original_option,
  `Provisional English display label` = d$English_display_label, check.names = FALSE),
  "Education definitions for translation review; no participant responses or subgroup counts.")
cat('<div class="review"><strong>Typography.</strong> Preferred font: ', esc(report$font$Preferred_font),
  '; installed: ', esc(report$font$Preferred_font_available), '. First fallback: ', esc(report$font$First_fallback),
  '; installed: ', esc(report$font$First_fallback_available), '. The chart renderer resolved regular text to ',
  esc(report$font$Resolved_regular_family), ' and bold text to ', esc(report$font$Resolved_bold_family),
  '. The same regular family is requested for this HTML report. Browser rendering should be rechecked on another computer.</div>', sep = "")
```

The blue, green and light-blue palette comes from the shared EUAqua theme and represents provisional visual approximations, not verified official brand colour codes. No appropriate official logo asset was found in the analysis workspace; no replacement logo has been created and no fonts have been downloaded.

Unconfirmed codes require survey/export evidence before future use. Education translations and the demographic disclosure rule require review. Correct the pilot-date metadata if the actual session differed from the schedule.

## Reproducibility

In RStudio, run `source("reports/render_pilot_report.R")` from the project root. It reruns the existing analysis, verifies tables against CSV outputs and renders numerical findings directly from aggregates.

Keep the HTML and its asset folders together under ignored `outputs/`. Numerical checks and R environment details are recorded in `verification.txt` and `session_info.txt`. Review browser print preview before printing or saving a PDF.
